master thesis

### 2018-03-24 Paper - Latex/Paper - Skizze - v4.tex
Paper - Discussion - Simulation

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities}
      As mentioned above, in an ancestral state reconstruction, transitions are minimized, with 
        transition probabilities playing a role. \\
      This subsection deals with these transition probabilities from free-living (hereinafter / as a 
        formula FL) to parasitic (hereinafter P) and vice versa: $\mathcal{P}(FL \rightarrow P)$, 
        $\mathcal{P}(P \rightarrow FL)$. \\
      Different parasite types have different transition probabilities. It is very difficult to make a 
        statement about these probabilities. It is generally assumed in this work that there are 40~\% 
        parasites and 60~\% free-livings which is based on the estimates by Windsor \cite{Windsor1998} 
        and $\mathcal{P}(FL \rightarrow P) > \mathcal{P}(P \rightarrow FL)$, because a reverse mutation 
        is usually less likely. \todo{This is discussed in section x of the discussion.} \\

@@

    A tree is needed to do a simulation of ancestral state reconstruction. It had to be decided whether 
      to take the real tree or simulate a tree. In this simulation, trees are created randomly, as 
      one can replicate a complete binary phylogentic tree. Thus, there is also the possibility to 
      simulate the multifurcation. \\
    To get a random binary tree, the Phylo package from biopython is used \cite{Cock2009}. They offer 
      a randomized function which returns a BaseTree\footnote{
        \hyperlink{https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
        {https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
      }. \\
    \todo{ref in die discussion über die randomized function? Diskutieren wir das?}

@@

      We examined the influence of the omitted data on the prediction. On average, about twice as many 
        leaf nodes are predicted differently. \color{red}The variance is very high. This means that 
        we have a high degree of dispersion and thus a stochastic situation exists. It also describes 
        the width of the present probability function. The standard deviation is about five times as 
        high as the number of omitted nodes, so the variability is quintupled. \normalfont Table 
        \ref{table:statistics cross validation} shows these results. \\
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance ($\sigma^2$) & $\sigma$ \\
            \hline \hline
            \multirow{3}{*}{\bfseries distance} & \bfseries all     & 0 & 3587.70 & 224.96 & 313650.61 & 560.05 \\
            & \bfseries leaf nodes                                  & 0 & 3021.12 & 208.69 & 248103.38 & 498.10 \\
            & \bfseries internal nodes                              & 0 & 566.58 & 16.28 & 4927.95 & 70.20 \\ \hline
            \multirow{3}{*}{\bfseries lost} & \bfseries all states  & 100 & 100 & 100 & 0 & 0 \\
            & \bfseries FL states                                   & 44 & 66 & 57.25 & 19.50 & 4.42 \\
            & \bfseries P states                                    & 34 & 56 & 42.75 & 19.50 & 4.42 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics about the leave-100-out cross validation \\
          The distance between original and new states was calculated using the Euclidean metric. This 
            was summed over all states / all leaf node states and all internal node states. \\
          The lower half of the table describes the distribution of the 'lost' states between parasites 
            (P) and free-livings (FL).}
        \label{table:statistics cross validation}
      \end{table}
    %---------------------------------------------------------------------------------------------------

      % \begin{table}
      %   \begin{center}
      %     \begin{tabular}{ |cl||r|r|r|r| }
      %       \hline
      %       & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance \\
      %       \hline \hline
      %       \rowcolor{green!50}distance & all   & 1 & 2 & 1.33 & 0.33 \\
      %       \rowcolor{orange!50}&               & 0 & 3587.70 & 217.94 & 273760.68 \\
      %       \rowcolor{green!50}& leaf nodes     & 1 & 2 & 1.33 & 0.33 \\
      %       \rowcolor{orange!50}&               & 0 & 3021.12 & 202.57 & 209274.86 \\
      %       \rowcolor{green!50}& internal nodes& 0 & 0 & 0.00 & 0.00 \\
      %       \rowcolor{orange!50}&               & 0 & 566.58 & 15.37 & 5684.00 \\ \hline
      %       \rowcolor{green!50}lost & FL states & 44 & 49 & 46.67 & 6.33 \\
      %       \rowcolor{orange!50}&               & 51 & 66 & 57.95 & 13.47 \\
      %       \rowcolor{green!50}& P states       & 51 & 56 & 53.33 & 6.33 \\
      %       \rowcolor{orange!50}&               & 34 & 49 & 42.05 & 13.47 \\
      %       \hline
      %     \end{tabular}
      %   \end{center}
      %   \caption{Statistics to cross validation \\
      %     - \colorbox{green!50}{Less free-livings - 3 examples} \\
      %     - \colorbox{orange!50}{Less parasites - 64 examples}}
      % \end{table}

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Discussion}

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Simulation}
    For fixed calculations we assume a distribution of 60\% free-living to 40\% parasites, 95\% 
      missing data and a multifuration rate of 95\%.
    How well does our simulation approach the real data situation? \\
    There are some points to discuss:
    \begin{itemize}
      \item How close is the randomized binary tree to a true phylogeny?
      \item The distribution of parasites to free-ranging species is a pure assumption. From different 
        sources about 40\% parasites are estimated, are these also beta-distributed?
      \item How well do the transition probabilities match parasitism? Depending on the type of 
        parasitism, this will certainly look different. In addition, it can be assumed that the 
        probability for losses is lower than for origins.
      \item The equal distribution of the multifurcation is a blank assumption. It is very likely to 
        be higher towards the root node because there is less information back in time. On the other 
        hand, there are many studies that have studied various species without considering phylogeny. 
        Which also creates a high number of children of a node directly at the leaf nodes.
    \end{itemize}
    In order to be able to observe these problems, we have programmed our simulation according to the 
      parameters. This allowed us to estimate the influence of some of these parameters.

>>>>

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities}
      As mentioned above, in an ancestral state reconstruction, transitions are minimized, with 
        transition probabilities playing a role. \\
      This subsection deals with these transition probabilities from free-living (hereinafter / as a 
        formula FL) to parasitic (hereinafter P) and vice versa: $\mathcal{P}(FL \rightarrow P)$, 
        $\mathcal{P}(P \rightarrow FL)$. \\
      Different parasite types have different transition probabilities. It is very difficult to make a 
        statement about these probabilities. It is generally assumed in this work that there are 40~\% 
        parasites and 60~\% free-livings which is based on the estimates by Windsor \cite{Windsor1998} 
        and $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$, as a result of the 
        diversity of parasites and the lack of general determinations for this. These parameters are 
        debated in subsection \ref{sec:discussion - transition probabilities} of the discussion. \\

@@

    A tree is needed to do a simulation of ancestral state reconstruction. It had to be decided whether 
      to take the real tree or simulate a tree. In this simulation, trees are created randomly, as 
      one can replicate a complete binary phylogentic tree. Thus, there is also the possibility to 
      simulate the multifurcation. \\
    To get a random binary tree, the Phylo package from biopython is used \cite{Cock2009}. They offer 
      a \textit{randomized()} function which returns a BaseTree\footnote{
        \hyperlink{https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
        {https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
      }. The credibility of this choice is examined in subsection 
        \ref{sec:discussion - simulate phylogeny} of the discussion.\\

@@

      We examined the influence of the omitted data on the prediction. On average, about twice as many 
        leaf nodes are predicted differently. \color{red}The variance is very high. This means that 
        we have a high degree of dispersion and thus a stochastic situation exists. It also describes 
        the width of the present probability function. The standard deviation is about five times as 
        high as the number of omitted nodes, so the variability is quintupled. \color{black} Table 
        \ref{table:statistics cross validation} shows these results. \\
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance ($\sigma^2$) & $\sigma$ \\
            \hline \hline
            \multirow{3}{*}{\bfseries distance} & \bfseries all     & 0 & 3587.70 & 224.96 & 313650.61 & 560.05 \\
            & \bfseries leaf nodes                                  & 0 & 3021.12 & 208.69 & 248103.38 & 498.10 \\
            & \bfseries internal nodes                              & 0 & 566.58 & 16.28 & 4927.95 & 70.20 \\ \hline
            \multirow{3}{*}{\bfseries lost} & \bfseries all states  & 100 & 100 & 100 & 0 & 0 \\
            & \bfseries FL states                                   & 44 & 66 & 57.25 & 19.50 & 4.42 \\
            & \bfseries P states                                    & 34 & 56 & 42.75 & 19.50 & 4.42 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics about the leave-100-out cross validation \\
          The distance between original and new states was calculated using the Euclidean metric. This 
            was summed over all states / all leaf node states and all internal node states. \\
          The lower half of the table describes the distribution of the 'lost' states between parasites 
            (P) and free-livings (FL).}
        \label{table:statistics cross validation}
      \end{table}

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Discussion}
  As in the previous chapters, the discussion is divided into three sections: input data, simulation 
    and reconstruction of the real data.

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Simulation}
    % For fixed calculations we assume a distribution of 60\% free-living to 40\% parasites, 95\% 
    %   missing data and a multifuration rate of 95\%.

    There are some points to discuss:
    \begin{itemize}
      \item How close is the randomized binary tree to a true phylogeny?
      \item The distribution of parasites to free-ranging species is a pure assumption. From different 
        sources about 40\% parasites are estimated, are these also beta-distributed?
      \item How well do the transition probabilities match parasitism? Depending on the type of 
        parasitism, this will certainly look different. In addition, it can be assumed that the 
        probability for losses is lower than for origins.
      \item The equal distribution of the multifurcation is a blank assumption. It is very likely to 
        be higher towards the root node because there is less information back in time. On the other 
        hand, there are many studies that have studied various species without considering phylogeny. 
        Which also creates a high number of children of a node directly at the leaf nodes.
    \end{itemize}
    In order to be able to observe these problems, we have programmed our simulation according to the 
      parameters. This allowed us to estimate the influence of some of these parameters.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities} \label{sec:discussion - transition probabilities}
      We assume that different parasite types have different transition probabilities. Establishing a 
        common distribution across the eukaryotic tree is therefore difficult. \\
      Based on the estimates of Windsor \cite{Windsor1998}, we have assumed a distribution of 40\% 
        parasites to 60\% free livings in this work. As a result of the diversity of parasites and the 
        lack of generalizations, we have generally stated that 
        $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$. But it is also reasonable to 
        assume that in general $\mathcal{P}(FL \rightarrow P) > \mathcal{P}(P \rightarrow FL)$, because 
        a reverse mutation is usually less likely. However, one would have to determine how much this 
        difference is and thus discuss another parameter. \\

      In the simulation, we tested different beta distributions with different thresholds and found that 
        this combination has a considerable influence on the predictive power of the algorithms. \\
      At this point it would be possible to test other distributions with equal threshold values or 
        different threshold values with equal distributions. \\
      However, if you choose a strong free-living distribution and balance it with the threshold to 
        achieve a 40:60 parasite free-living distribution, for example, this could be very 
        computationally expensive. \\

      With these further simulations you could find out if the issue of distributions plays a big role. 
        Conversely, one could estimate possible distributions based on the data location in the tree. 
        However, this is likely to be very difficult given the poor data. For this reason, we have 
        decided in this work to accept very general values and not to speculate much. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Simulation of a phylogeny} \label{sec:discussion - simulate phylogeny}
      In our simulation, we start with the simulation of a phylogeny and then depict our data situation. \\
      For this first step we use the \textit{randomized()} function of the phylo package from biopython
        \cite{Cock2009}. The question is how close is such a randomized tree to a phylogeny. \\

      The problems that can arise are that different species develop at different rates. This means that 
        there could be sections in the tree with many branches (for example, subtrees of unicellular 
        species), ie with high depths and opposite sections (of very complex species, for example). It 
        can thus be assumed that the tree is not balanced. \\
      For more precise statements, one would have to take deeper undercurrents, which was not possible 
        in the context of this work.
